---
title: "HW5_Q4"
author: "Ethan M. Lotz"
date: "November 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Biobase)
library(marray)
library(affy)
```

## Question 4
The three custom spotted array data files GMU1420.txt, GMU1421.txt, and GMU1422.txt are three technical
replicates whereby the Stratagene Universal Human RNA was hybridized. Read the data into the R
programming environment. Use code from in-class exercises, change names of the columns containing
Red/Green intensities as to NOT contain spaces, like: name.Gf = “ch1Intensity”

```{r}
setwd("C:/Users/lotzem/Desktop/School/Classes/BIOS 567/Homework/Homework 5/hw5/q4/")
file <- dir(pattern = "txt")[1:3]
file
layout<-read.marrayLayout(fname=file[1], ngr=12, ngc=4, nsr=21, nsc=21, skip=21211)	
info<-read.marrayInfo(fname=file[1], info.id=c(1,6), labels=6, skip=21211)
raw.data <- read.marrayRaw(file, 
                           name.Gf="ch1Intensity", name.Gb="ch1Background", 
                           name.Rf="ch2Intensity", name.Rb="ch2Background", 
                           layout=layout, gnames=info, skip=21211)
```

a.) Produce all chip-specific MA plots.
```{r}
maPlot(raw.data[,1],x="maA",y="maM",z=NULL,lines.func=NULL)
maPlot(raw.data[,2],x="maA",y="maM",z=NULL,lines.func=NULL)
maPlot(raw.data[,3],x="maA",y="maM",z=NULL,lines.func=NULL)
```


b.) For GMU1420.txt, produce a scatterplot of the Green foreground by Green background, as well as a plot of Red foreground by Red background.
```{r}
plot(log2(raw.data@maGf[,1]),log2(raw.data@maGb[,1]), ylab="Green Foreground", xlab="Green Background")
plot(log2(raw.data@maRf[,1]),log2(raw.data@maRb[,1]), ylab="Red Foreground", xlab="Red Background")

```

What, if anything, do these plots suggest?

The range of the red channel is larger than the green channel, but both graphs indicate a constant linear relationship suggesting minimal dye bias.")

c.) Apply constant global normalization to the raw data; plot the MA plots and boxplots by print-tip for this post normalized data.
```{r}
raw.data.norm.m <- maNorm(raw.data, norm="median")

maPlot(raw.data.norm.m[,1],x="maA",y="maM",z=NULL,lines.func=NULL, main="Constant Global Normalization-1")
boxplot(raw.data.norm.m[,1], xvar="maPrintTip", yvar="maM", main="Constant Global Normalization-1")

maPlot(raw.data.norm.m[,2],x="maA",y="maM",z=NULL,lines.func=NULL, main="Constant Global Normalization-2")
boxplot(raw.data.norm.m[,2], xvar="maPrintTip", yvar="maM", main="Constant Global Normalization-2")

maPlot(raw.data.norm.m[,3],x="maA",y="maM",z=NULL,lines.func=NULL, main="Constant Global Normalization-3")
maBoxplot(raw.data.norm.m[,3], xvar="maPrintTip", yvar="maM", main="Constant Global Normalization-3")
```

d.) Apply global loess normalization to the raw data; plot the MA plots and boxplots by print-tip for this post normalized data.
```{r}
raw.data.norm.l <- maNorm(raw.data, norm="loess")

maPlot(raw.data.norm.l[,1],x="maA",y="maM",z=NULL,lines.func=NULL, main="Global Loess Normalization-1")
boxplot(raw.data.norm.l[,1], xvar="maPrintTip", yvar="maM", main="Global Loess Normalization-1")

maPlot(raw.data.norm.l[,2],x="maA",y="maM",z=NULL,lines.func=NULL, main="Global Loess Normalization-2")
boxplot(raw.data.norm.l[,2], xvar="maPrintTip", yvar="maM", main="Global Loess Normalization-2")

maPlot(raw.data.norm.l[,3],x="maA",y="maM",z=NULL,lines.func=NULL, main="Global Loess Normalization-3")
boxplot(raw.data.norm.l[,3], xvar="maPrintTip", yvar="maM", main="Global Loess Normalization-3")
```

e.) Apply print tip loess normalization to the raw data; plot the MA plots and boxplots by print tip for this post normalized data.
```{r}
raw.data.norm.p <- maNorm(raw.data, norm="printTipLoess")

maPlot(raw.data.norm.p[,1],x="maA",y="maM",z=NULL,lines.func=NULL, main="Print-Tip Loess Normalization-1")
boxplot(raw.data.norm.p[,1], xvar="maPrintTip", yvar="maM", main="Print-Tip Loess Normalization-1")

maPlot(raw.data.norm.p[,2],x="maA",y="maM",z=NULL,lines.func=NULL, main="Print-Tip Loess Normalization-2")
boxplot(raw.data.norm.p[,2], xvar="maPrintTip", yvar="maM", main="Print-Tip Loess Normalization-2")

maPlot(raw.data.norm.p[,3],x="maA",y="maM",z=NULL,lines.func=NULL, main="Print-Tip Loess Normalization-3")
boxplot(raw.data.norm.p[,3], xvar="maPrintTip", yvar="maM", main="Print-Tip Loess Normalization-3")

```

f.) Write a brief paragraph comparing and contrasting these normalization methods applied to this set of custom spotted array data.

All three methods do a fairly decent job of normalizing the data. Constant global normalization centers the MA plots around zero but does not do a good job of adjusting the skewness of the data. In addition the distribution of the box plots still vary quite a bit. Global Loess normalization does a much better job of eliminating the skewness of the data in the MA plots as well as centering their distribution around zero. However, the boxplots still vary. The Print-Tip Loess normalization does as well of a job as the global Loess normalization technique regarding the MA plots; however, the box plot distributions are much more compact and vary much less compared to the other normalization methods.

