---
title: "HW5_Q5"
author: "Ethan M. Lotz"
date: "November 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(affy)
library(Biobase)
library(geneplotter)
library(genefilter)
```

## Question 5

Read the posted *.CEL files (Affymetrix GeneChip data) into the R programming environment. These two GeneChips were each hybridized with the same RNA, hence, they are technical replicates.
```{r}
setwd("C:/Users/lotzem/Desktop/School/Classes/BIOS 567/Homework/Homework 5/hw5/q5/")
list.celfiles()
GeneChips <- read.affybatch(filenames = list.celfiles(pattern = "Ref*"))
pData(GeneChips)
GeneChips
```

a.) Produce a boxplot of the PM and MM intensities signals by GeneChip.
```{r}
#Chip 1
boxplot(log2(pm(GeneChips)[, 1]), log2(mm(GeneChips)[, 1]), main="PM and MM 1st Chip")

#Chip 2
boxplot(log2(pm(GeneChips)[, 2]), log2(mm(GeneChips)[, 2]), main="PM and MM 2nd Chip")

```

b.) Produce the MA plots using only the PM intensities for these GeneChips.
```{r}
ma.plot <- function(x, y) {
  M <- log(x, 2) - log(y, 2)
  A <- (log(x, 2) + log(y, 2))/2
  plot(A, M)
}

ma.plot(pm(GeneChips)[, 1], pm(GeneChips)[, 2])
```

c.) Apply quantile normalization to this set of data; plot the resulting MA plot for the post-normalized PM data for these GeneChips.
```{r}
GeneChips.norm.q <- normalize(GeneChips, method = "quantiles")

ma.plot(pm(GeneChips.norm.q)[, 1], pm(GeneChips.norm.q)[, 2])
```

d.) Apply invariant set normalization and plot the resulting MA plot for the post- normalized PM data for these GeneChips.
```{r}
GeneChips.norm.i <- normalize(GeneChips, method = "invariantset")
ma.plot(pm(GeneChips.norm.i)[, 1], pm(GeneChips.norm.i)[, 2])
```

e.) Apply loess normalization and plot the resulting MA plot for the post- normalized PM data for these GeneChips.
```{r}
GeneChips.norm.l <- normalize(GeneChips, method = "loess")
ma.plot(pm(GeneChips.norm.l)[, 1], pm(GeneChips.norm.l)[, 2])
```

